---
title: "exercise-11"
format: html
author: Qiaoyun Ai
---

#### Hide unnecessary message

```{r}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

#### **Preliminaries**

```{r}
library(tidyverse)
library(skimr)
library(naniar)
library(broom)
library(MuMIn)

f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f, col_names = TRUE)
names (d)
```

#### Do a bit of exploratory data analysis with this dataset, e.g., using the {skimr} package.

```{r}
skim(d)
```

##### Q: Which of the variables are categorical and which are numeric?

##### A: The variables order, family, Genus, and species are categorical, the rest are numeric.

#### Step 1: Replace all values of -999 (the authors’ code for missing data) with `NA`.

```{r}
d <- d |> replace_with_na_all(condition = ~.x == -999)
```

#### Step 2

```{r}
# Drop the variables litter size and refs.
d <- d |>
    select(-c("litter size", "refs"))

# Rename the variable max. life(mo) to maxlife(mo) & wean mass(g) to weanmass(g).
d <- d |>
    rename(`maxlife(mo)` = `max. life(mo)`) |>
    rename(`weanmass(g)` = `wean mass(g)`)
```

#### Step 3

#### Log transform all of the other numeric variables.

```{r}
d <- d |>
    mutate(across(where(is.numeric), log))

summary(d)
```

#### Step 4

#### Regress the (now log transformed) *age* \[**gestation(mo)**, **weaning(mo)**, **AFR(mo)** (i.e., age at first reproduction), and **maxlife(mo)** (i.e., maximum lifespan)\] and *mass* \[**newborn(g)** and **weanmass(g)**\] variables on (now log transformed) overall body **mass(g)** and add the residuals to the dataframe as new variables \[**relGest**, **relWean**, **relAFR**, **relLife**, **relNewbornMass**, and **relWeaningMass**\].

```{r}
d$relGest <- resid(lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude))
d$relWean <- resid(lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude))
d$relAFR <- resid(lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude))
d$relLife <- resid(lm(`maxlife(mo)` ~ `mass(g)`, data = d, na.action = na.exclude))
d$relNewbornMass <- resid(lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude))
d$relWeaningMass <- resid(lm(`weanmass(g)` ~ `mass(g)`, data = d, na.action = na.exclude))
```

#### Step 5

#### Plot residuals of max lifespan (**relLife**) in relation to **Order**.

```{r}
p1 <- ggplot(data = d, aes(x = order, y = relLife)) + geom_boxplot(na.rm = TRUE) +
    geom_jitter(na.rm = TRUE, alpha = 0.1) + theme(axis.text.x = element_text(angle = 90))
p1
```

##### Q: Which mammalian orders have the highest residual lifespan?

##### A: Order Primates has the highest residual lifespan.

#### Plot residuals of newborn mass (**relNewbornMass**) in relation to **Order**.

```{r}
p2 <- ggplot(data = d, aes(x = order, y = relNewbornMass)) + geom_boxplot(na.rm = TRUE) +
    geom_jitter(na.rm = TRUE, alpha = 0.1) + theme(axis.text.x = element_text(angle = 90))
p2
```

##### Q: Which mammalian orders have the have highest residual newborn mass?

##### A: Orders Macroscelidae and Cetacea have the highest residual newborn mass.

#### Plot residuals of weaning mass (**relWeaningMass**) in relation to **Order**.

```{r}
p3 <- ggplot(data = d, aes(x = order, y = relWeaningMass)) + geom_boxplot(na.rm = TRUE) +
    geom_jitter(na.rm = TRUE, alpha = 0.1) + theme(axis.text.x = element_text(angle = 90))
p3
```

##### Q: Which mammalian orders have the have highest residual weaning mass?

##### A: Order Perrisodactyla has the highest residual weaning mass.

#### Step 6

#### Run models and a model selection process to evaluate what (now log transformed) variables best predict each of the two response variables, **maxlife(mo)** and **AFR(mo)**, from the set of the following predictors: **gestation(mo)**, **newborn(g)**, **weaning(mo)**, **weanmass(g)**, **litters/year**, and overall body **mass(g)**.

#### For maxLife

```{r}
# maxLife
# maxlife dataset
d_life <- d %>%
  drop_na(`maxlife(mo)`, `gestation(mo)`, `newborn(g)`,
          `weaning(mo)`, `weanmass(g)`,
          `litters/year`, `mass(g)`)

# Full and null models
ML_full <- lm(`maxlife(mo)` ~ `gestation(mo)` + `newborn(g)` +
                `weaning(mo)` + `weanmass(g)` +
                `litters/year` + `mass(g)`,
              data = d_life, na.action = na.fail)
summary(ML_full)

ML_null <- lm(`maxlife(mo)` ~ 1, data = d_life, na.action = na.fail)

# Backward selection
drop1(ML_full, test = "F")
b2 <- update(ML_full, . ~ . - `weanmass(g)`)
drop1(b2, test = "F")
b3 <- update(b2, . ~ . - `newborn(g)`)
drop1(b3, test = "F")

# Forward selection
add1(ML_null,
     . ~ . + `gestation(mo)` + `newborn(g)` +
       `weaning(mo)` + `weanmass(g)` +
       `litters/year` + `mass(g)`,
     test = "F")
f2 <- update(ML_null, . ~ . + `gestation(mo)`)
add1(f2,
     . ~ . + `newborn(g)` + `weaning(mo)` +
       `weanmass(g)` + `litters/year` + `mass(g)`,
     test = "F")
f3 <- update(f2, . ~ . + `litters/year`)
add1(f3,
     . ~ . + `newborn(g)` + `weaning(mo)` +
       `weanmass(g)` + `mass(g)`,
     test = "F")
f4 <- update(f3, . ~ . + `mass(g)`)
add1(f4,
     . ~ . + `newborn(g)` + `weaning(mo)` + `weanmass(g)`,
     test = "F")
f5 <- update(f4, . ~ . + `weaning(mo)`)
add1(f5, . ~ . + `newborn(g)` + `weanmass(g)`, test = "F")

# using AICc
ML_mods <- dredge(ML_full)
head(coef(ML_mods), 10)

(ML_mods.avg <- summary(model.avg(ML_mods, subset = delta < 4)))

# Calculate and plot the model-averaged coefficients and their CIs across this set of top models
confint(ML_mods.avg)
plot(ML_mods.avg, full = TRUE, intercept = FALSE)
```

#### For AFR

```{r}
# AFR
# AFR dataset
d_afr <- d %>%
  drop_na(`AFR(mo)`, `gestation(mo)`, `newborn(g)`,
          `weaning(mo)`, `weanmass(g)`,
          `litters/year`, `mass(g)`)

# Full and null models
AFR_full <- lm(`AFR(mo)` ~ `gestation(mo)` + `newborn(g)` +
                 `weaning(mo)` + `weanmass(g)` +
                 `litters/year` + `mass(g)`,
               data = d_afr, na.action = na.fail)
summary(AFR_full)

AFR_null <- lm(`AFR(mo)` ~ 1, data = d_afr, na.action = na.fail)

# Backward selection
drop1(AFR_full, test = "F")
b2 <- update(AFR_full, . ~ . - `weanmass(g)`)
drop1(b2, test = "F")
b3 <- update(b2, . ~ . - `newborn(g)`)
drop1(b3, test = "F")

# Forward selection
add1(AFR_null,
     . ~ . + `gestation(mo)` + `newborn(g)` +
       `weaning(mo)` + `weanmass(g)` +
       `litters/year` + `mass(g)`,
     test = "F")
f2 <- update(AFR_null, . ~ . + `litters/year`)
add1(f2,
     . ~ . + `gestation(mo)` + `newborn(g)` +
       `weaning(mo)` + `weanmass(g)` + `mass(g)`,
     test = "F")
f3 <- update(f2, . ~ . + `gestation(mo)`)
add1(f3,
     . ~ . + `newborn(g)` + `weaning(mo)` +
       `weanmass(g)` + `mass(g)`,
     test = "F")
f4 <- update(f3, . ~ . + `weanmass(g)`)
add1(f4,
     . ~ . + `newborn(g)` + `weaning(mo)` + `mass(g)`,
     test = "F")
f5 <- update(f4, . ~ . + `weaning(mo)`)
add1(f5, . ~ . + `newborn(g)` + `mass(g)`, test = "F")

# using AICc
AFR_mods <- dredge(AFR_full)
head(coef(AFR_mods), 10)

(AFR_mods.avg <- summary(model.avg(AFR_mods, subset = delta < 4)))

# Calculate and plot the model-averaged coefficients and their CIs across this set of top models
confint(AFR_mods.avg)
plot(AFR_mods.avg, full = TRUE, intercept = FALSE)
```

##### Q1: For each of the two response variables, indicate what is the best model overall based on AICc and how many models have a delta AICc of 4 or less?

##### A: The best model for maxlife(mo) is maxlife(mo) \~ gestation(mo) + litters/year + mass(g) + weaning(mo). Five models have delta AICc of 4 or less

##### The best model for AFR(mo) is AFR(mo) \~ gestation(mo) + litters/year + mass(g) + weaning(mo). Seven models have delta AICc of 4 or less

##### Q2: What variables, if any, appear in all of this set of “top” models?

##### A: For maxlife(mo), the variables present in all top models are: gestation(mo), litters/year, mass(g), and weaning(mo)

##### For AFR(mo), the variables present in all top models are: gestation(mo), litters/year, and weaning(mo)
